Convergence rate of MAP estimates \\ for the exponential family

### 2021-10-09 main.tex
Close self-concordance discussion

\section{Stochastic Bregman Proximal Point}
\subsection{Correspondence with MAP}
For a dataset of independent samples $\cD_n = \{X_1,\dots,X_n\}$, the posterior can be seen iteratively
\begin{align}
	p(\natp |\cD_n)
	=p(\natp|X_n,\cD_{n-1})
	\propto p(X_n|\natp,\cD_{n-1}) p(\natp|\cD_{n-1})
	=p(X_n|\natp)p(\natp |\cD_{n-1}) \; .
\end{align}
For the MAP of the exponential family with a Bregman conjugate prior, the posterior takes the form
\begin{align}
	- \log p(\natp | \cD_n) 
	&= \sum_{i=1}^n -\log p(X_i | \natp) - \log p(\natp) + \cst\\
	&= \sum_i - \langle T(X_i) ; \natp \rangle  + n \logpart(\natp) + n_0 \bregman(\natp ; \natp_0) +\cst \\
	&= (n+n_0)\logpart(\natp)  - \langle n_0 \mu_0 + \sum_i T(X_i) ; \natp \rangle + \cst \\
	&= (n+n_0) \bregman(\natp; \hat \natp_n^\text{MAP}) + \cst \; ,
\end{align}
so we immediately get the recursion
\begin{align}
    \hat \natp_n^\text{MAP}
    &= \argmin_\natp - \log p(\natp | \cD_n) \\
    &= \argmin_\natp - \log p(X_n|\natp) - \log p(\natp | \cD_{n-1}) \\
    &= \argmin_\natp f(\natp; X_{n}) + (n+n_0) \bregman(\natp; \hat \natp_{n-1}^\text{MAP})  \; .
    \label{eq:SBPP}
\end{align} 
Coincidentally, this is exactly the formula of a Stochastic Bregman Proximal Point update on the function $f(\natp) = \E_X[-\log p(X|\natp)]$ , with step-size $\lr_n = \inv{n+n_0}$. As a reminder this algorithm is defined with the update
\begin{align}
	 \hat \natp_n
    &= \argmin_\natp \lr_n f(\natp; X_{n}) + \bregman(\natp; \hat \natp_{n-1}) \\
    \iff \nabla \logpart(\natp_{n-1}) &= \nabla\logpart(\natp_n) + \lr_t \nabla f(\natp_n ; X_t) \; .
\end{align}
In words, if it converges to $\natp^*$, it can be seen as a backward mirror ascent starting from $\natp^*$.

@@

\paragraph{Notations}
\begin{itemize}
    \item $\|\cdot\|_x = \sqrt{ \langle \nabla^2A^*(x)\cdot,\, \cdot \rangle }$
    \item $\|\cdot\|_x^*= \sqrt{ \langle [\nabla^2A^*(x)]^{-1}\cdot,\, \cdot \rangle }$
    \item $\lambda(x) = \|(A^*)'(x)\|_x^*$ (Newton decrements)
    \item $\omega = t-\ln (1+t)$
    \item $\omega^* = -t-\ln(1-t)$
\end{itemize}
Note that $\omega^*$ is convex and monotonne.

@@

The last proposition implies that, when $\|y-x\|_x < 1$,
\[
    \mathcal{B}_{A^*}(y,x) \leq \omega^*(\|y-x\|_x).
\]
Therefore,
\[
    \mathcal{B}_{A^*}(\mu^*,\mu_n) \leq \omega^*(\|\mu^*-\mu_n\|_{\mu_n}) .
\]
then, if $\|\mu^*-\mu_n\|_{\mu^*} < 1$, we have
\[
    \mathcal{B}_{A^*}(\mu^*,\mu_n) \leq \omega^*\left(\frac{\|\mu^*-\mu_n\|_{\mu^*}}{1-\|\mu^*-\mu_n\|_{\mu^*}}\right) .
\]
Assume $\|\mu^*-\mu_n\|_{\mu^*} < c$,
\[
    \mathcal{B}_{A^*}(\mu^*,\mu_n) \leq \omega^*\left(\frac{\|\mu^*-\mu_n\|_{\mu^*}}{1-c}\right) .
\]

>>>>

\section{Stochastic Bregman Proximal Point}
\subsection{Correspondence with MAP}
For a dataset of independent samples $\cD_n = \{X_1,\dots,X_n\}$, the posterior can be seen iteratively
\begin{align}
	p(\natp |\cD_n)
	=p(\natp|X_n,\cD_{n-1})
	\propto p(X_n|\natp,\cD_{n-1}) p(\natp|\cD_{n-1})
	=p(X_n|\natp)p(\natp |\cD_{n-1}) \; .
\end{align}
For the MAP of the exponential family with a Bregman conjugate prior, the posterior takes the form
\begin{align}
	- \log p(\natp | \cD_n) 
	&= \sum_{i=1}^n -\log p(X_i | \natp) - \log p(\natp) + \cst\\
	&= \sum_i - \langle T(X_i) ; \natp \rangle  + n \logpart(\natp) + n_0 \bregman(\natp ; \natp_0) +\cst \\
	&= (n+n_0)\logpart(\natp)  - \langle n_0 \mu_0 + \sum_i T(X_i) ; \natp \rangle + \cst \\
	&= (n+n_0) \bregman(\natp; \hat \natp_n^\text{MAP}) + \cst \; ,
\end{align}
so we immediately get the recursion
\begin{align}
    \hat \natp_n^\text{MAP}
    &= \argmin_\natp - \log p(\natp | \cD_n) \\
    &= \argmin_\natp - \log p(X_n|\natp) - \log p(\natp | \cD_{n-1}) \\
    &= \argmin_\natp f(\natp; X_{n}) + (n+n_0 - 1) \bregman(\natp; \hat \natp_{n-1}^\text{MAP})  \; .
    \label{eq:SBPP}
\end{align} 
Coincidentally, this is exactly the formula of a Stochastic Bregman Proximal Point update on the function $f(\natp) = \E_X[-\log p(X|\natp)]$ , with step-size $\lr_n = \inv{n+n_0 -1}$. As a reminder this algorithm is defined with the update
\begin{align}
	 \hat \natp_n
    &= \argmin_\natp \lr_n f(\natp; X_{n}) + \bregman(\natp; \hat \natp_{n-1}) \\
    \iff \nabla \logpart(\natp_{n-1}) &= \nabla\logpart(\natp_n) + \lr_t \nabla f(\natp_n ; X_t) \; .
\end{align}
In words, if it converges to $\natp^*$, it can be seen as a backward mirror ascent starting from $\natp^*$.

@@

\paragraph{Notations}
\begin{itemize}
    \item $\|\cdot\|_x = \sqrt{ \langle \nabla^2A^*(x)\cdot,\, \cdot \rangle }$
    \item $\|\cdot\|_x^*= \sqrt{ \langle [\nabla^2A^*(x)]^{-1}\cdot,\, \cdot \rangle }$
    \item $\lambda(x) = \|(A^*)'(x)\|_x^*$ (Newton decrements)
    \item $\omega(t) = t-\ln (1+t)$
    \item $\omega^*(t) = -t-\ln(1-t)$
\end{itemize}
Note that $\omega^*$ is convex and monotone, and $\omega^*(t) \leq t^2, \forall t\leq 0.68$.

@@

The last proposition implies that, when $\|y-x\|_x < 1$,
\[
    \mathcal{B}_{A^*}(y,x) \leq \omega^*(\|y-x\|_x).
\]
Therefore,
\[
    \mathcal{B}_{A^*}(\mu^*,\mu_n) \leq \omega^*(\|\mu^*-\mu_n\|_{\mu_n}) .
\]
then, if $\|\mu^*-\mu_n\|_{\mu^*} < 1$, we have
\[
    \mathcal{B}_{A^*}(\mu^*,\mu_n) \leq \omega^*\left(\frac{\|\mu^*-\mu_n\|_{\mu^*}}{1-\|\mu^*-\mu_n\|_{\mu^*}}\right) .
\]
Assume $\|\mu^*-\mu_n\|_{\mu^*} < 0.21$, then
\[
    \bregmanconj(\mu^*,\mu_n) \leq \|\mu^*-\mu_n\|_{\mu^*}^2
\]
so we have a local quadratic behavior. To put into euclide's world. Problem is it requires global self-concordance.
